Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference Imaging Photometric Survey.}

### 2021-04-21 cluster_paper/ms.tex
update ms
Co-authored-by: Jason Curtis <jasonleecurtis@gmail.com>

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrp}{(\mathrm{Bp}-\mathrm{Rp})_0}

@@

\keywords{
	stellar ages (1581),
  stellar associations (1582),
  open clusters (1160),
  stellar dynamics (1596),
  stellar rotation (1629)
}

@@

Although it has been possible to detect dispersed stellar associations
for a long time, Gaia is now enabling major advances \citep[{\it
e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by the most recent studies can be summarized as
follows.  On one end are coeval groups with no discernable cores ({\it
e.g.}, Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here,
we focus on an intermediate regime: low-density halos associated with
a single overdensity, typically an open cluster that was known before
Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos may actually be tidal tails, as have been
reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
Berenices \citep{tang_comaber_2019}.

@@

As part of a Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we have recently been making TESS light
curves of stars reported to be members of coeval populations and
searching them for planets \citep{bouma_cluster_2020}.  Our analysis
of TESS Sectors 1-13 yielded light curves of \numsouthernuniqlcs\
candidate cluster members in the Southern Ecliptic hemisphere, which
are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100--200\,Myr) and
sufficiently nearby ($d\approx400$\,pc) to facilitate photometric
rotation measurements using TESS.  Its halo was recently reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc.  We want to know: is this halo real
or is it just a collection of interlopers, foreground and background
stars?  To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available
\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
accuracy of the Gaia-derived results are the current state of the art.
We therefore adopted what we viewed as the most important Gaia-based
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of these samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}.
%While we considered performing our own clustering analysis on the
%Gaia data, such an effort would in effect only replicate the work of
%these investigators.  We opt to instead use their studies as a starting
%point.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2);
    these model values are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$
color.  Stars in the core appear consistent with having a fixed age
and metallicity, and varying mass.  The halo stars show a similar
sequence, though with greater scatter.  One possible explanation for
the additional scatter is that the halo is more contaminated by
field stars.  For instance, $\approx5$ red giants in the halo must be
field interlopers, because their isochronally implied ages would be
inconsistent with that of the cluster.

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.

To determine the reddening correction across \cn, we queried the 2MASS
DUST service\footnote{
  \url{http://irsa.ipac.caltech.edu/applications/DUST}; query
  performed using the \texttt{astrobase.services.dust} module
  \citep{bhatti_astrobase_2018}.  } and retrieved the total
line-of-sight extinction parameters at the positions of each \cn\ 
member.  The mean and standard deviation of the
$E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}:
$E(\mathrm{B}-\mathrm{V})_{\rm S}=0.206\pm0.039$.
\citet{bonifacio_search_2000} found however that the
\citet{schlegel_maps_1998} maps overestimate the reddening values when
the color excess exceeds about 0.10\,mag. We therefore applied the
correction proposed by \citet{bonifacio_search_2000}, and additionally
corrected for the mean cluster distance by a factor
$\exp(-|d\sin(b)|/h)$ where $d$ is the average cluster distance,
$b$ is the average galactic latitude, and $h$ is the scale height of
the galactic disk, assumed to be 125\,pc.  This yielded our adopted
$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$, where the uncertainty
reflects the standard deviation of the individual
\citet{schlegel_maps_1998} values.  Finally, to convert to Gaia
colors we used the calibration of \citet{stassun_TIC8_2019}, namely
$E(\mathrm{Bp}-\mathrm{Rp})=1.31 E(\mathrm{B}-\mathrm{V})$ and
$A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded
$E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$, which is used in the
plots.  To redden the isochrones, we assumed $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.

For the metallicity, we considered a range of super and sub-solar
metallicities to fit as much of the locus from
$0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as possible, and settled on the
slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
Sub-solar metallicities led to model predictions too blue along the
main sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees
with the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We
caution though that other sub-solar metallicities have been reported
\citep{bailey_rv_2018}, and note that the cluster metallicity and
reddening are degenerate; if the reddening is lowered, the implied
metallicity will increase.  While a detailed redetermination of
these parameters from the Gaia photometry is beyond our scope, our
adopted values for both the reddening and metallicity are within the
range of previously reported values in the literature.

Overall, the data and models agree for masses above
$\approx$0.45\,$M_\odot$.  Below this mass, the data and models
diverge at colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in
the sense that the model isochrones have lower luminosities and bluer
colors than the observations.\footnote{The 100 Myr isochrone does
overlap well with the lower main sequence, but it fails to fit the
upper main sequence and red giants.}  The MIST isochrones showed a
comparable disagreement \citep{choi_mesa_2016}.  Proposed explanations
for the discrepancy between the models and observations have include
starspots and incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximum
sensitivity to age.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3a.pdf}
			\includegraphics[width=0.49\textwidth]{f3b.pdf}
		}

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for
  \cn.} {\it Top row}: \cn\ split into its core and halo.  Sets
  $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo
  successive stages of automated cleaning, (see
  Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrp$$<$1.2.  From $1.2<\bpmrp<1.7$, some NGC\,2516
  members have longer rotation periods than Pleaides members.  {\it
  Bottom row}: \cn\ colored by binarity indicators.  Sources with
  $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\% of
  stars).  Overplotted in orange are sources at least 0.3\,mag above
  an empirically fit cluster isochrone (photometric binaries; 20\% of
  stars).
  Rotation periods for comparison field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $\bpmrp$ of 2.2, or a spectral type
of $\approx$M2V.

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.  The
``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrp<1.2$.  The implications of this comparison for the age of
\cn\ are discussed in Section~\ref{disc:absage}.

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr},
fixing the nodes by hand.  We then labeled any points brighter than
0.3\,mag above the cluster sequence as photometric binaries.  This
included 20\% of the overall cluster sample.  Many of the resulting
candidate binaries, though not all of them, appear below the slow
rotation sequence.  In set $\mathcal{B}$, over the color range where
the slow and fast sequence are present ($0.5<\bpmrp<1.3$), there are
281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
281 stars into ``fast'' and ``slow'' sequences by eye,  the fraction
of stars showing signs of binarity in the slow and fast subsamples are
30\% (63/210) and 56\% (40/71), respectively.  We speculate on
possible explanations in Section~\ref{disc:rotn}.

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrp<1.2$ for which our
TESS pipeline succeeding in making light curves.  In other words, the
stars in the base sample are those for which we could have plausibly
detected a rotation period.  This color range is preferred because the
the slow sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.  With $\approx$3.5\,pc, the period detection rate of
61\% is relatively low, due to the combined effects of crowding and
saturation.

@@

    members have detected rotation periods.  This fraction decreases

>>>>

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
%\newcommand{\bpmrpo}{(\bpmrp)_0}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

@@

\keywords{
	stellar ages (1581),
  stellar associations (1582),
  open clusters (1160),
  stellar kinematics (1608),
  stellar rotation (1629)
}

@@

Although it has been possible to detect dispersed stellar associations
for a long time, Gaia is now enabling major advances \citep[{\it
e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by the most recent studies can be summarized as
follows.  On one end are coeval groups with no discernable cores ({\it
e.g.}, Psc-Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here,
we focus on an intermediate regime: low-density halos associated with
a single overdensity, typically an open cluster that was known before
Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos may actually be tidal tails, as have been
reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
Berenices \citep{tang_comaber_2019,furnkranz_2019}.

@@

As part of a Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we have recently been making TESS light
curves of stars reported to be members of coeval populations and
searching them for planets \citep{bouma_cluster_2020}.  Our analysis
of TESS Sectors 1-13 yielded light curves of \numsouthernuniqlcs\
candidate cluster members in the Southern Ecliptic hemisphere, which
are available on
MAST.\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100-200\,Myr) and
sufficiently nearby ($d\approx400$\,pc) to facilitate photometric
rotation measurements using TESS.  Its halo was recently reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc.  We want to know: is this halo real
or is it just a collection of interlopers, foreground and background
stars?  To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available
\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
accuracy of the Gaia-derived results are the current state of the art.
We therefore adopted what we viewed as the most important Gaia-based
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of these samples is
available online.\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}
%While we considered performing our own clustering analysis on the
%Gaia data, such an effort would in effect only replicate the work of
%these investigators.  We opt to instead use their studies as a starting
%point.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Gaia EDR3 photometry for the core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models (and to give the approximate spectral
    types).  The opacity of the M-dwarf model atmospheres is less
    well-calibrated for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V);
    these model values are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $\bpmrp$ color.  Spectral
types are interpolated from the \citet{pecaut_mamajek_2013} table,
after accounting for reddening as described
below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2021.03.02.} Stars in the core appear consistent with having a
fixed age and metallicity, and varying mass.  The halo stars show a
similar sequence, though with greater scatter.  One possible
explanation for the additional scatter is that the halo is more
contaminated by field stars.  For instance, $\approx5$ red giants in
the halo must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface,\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26} \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.

To determine the reddening correction across \cn, we queried the 2MASS
DUST service\footnote{
  \url{http://irsa.ipac.caltech.edu/applications/DUST}; query
  performed using the \texttt{astrobase.services.dust} module
  \citep{bhatti_astrobase_2018}.  } and retrieved the total
line-of-sight extinction parameters at the positions of each \cn\ 
member.  The mean and standard deviation of the
$E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}:
$E(\mathrm{B}-\mathrm{V})_{\rm S}=0.206\pm0.039$.
\citet{bonifacio_search_2000} found however that the
\citet{schlegel_maps_1998} maps overestimate the reddening values when
the color excess exceeds about 0.10\,mag. We therefore applied the
correction proposed by \citet{bonifacio_search_2000}, and additionally
corrected for the mean cluster distance by a factor
$\exp(-|d\sin(b)|/h)$ where $d$ is the average cluster distance,
$b$ is the average galactic latitude, and $h$ is the scale height of
the galactic disk, assumed to be 125\,pc.  This yielded our adopted
$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$, where the uncertainty
reflects the standard deviation of the individual
\citet{schlegel_maps_1998} values.  Finally, to convert to Gaia
colors we used the calibration of \citet{stassun_TIC8_2019}, namely
$E(\bpmrp)=1.31 E(\mathrm{B}-\mathrm{V})$ and
$A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded
$E(\bpmrp)=0.134\pm0.025$, which is used in the
plots.  To redden the isochrones, we assumed $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.

For the metallicity, we considered a range of super and sub-solar
metallicities to fit as much of the locus from
$0.5<\bpmrp<1.5$ as possible, and settled on the
slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
Sub-solar metallicities led to model predictions too blue along the
main sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees
with the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We
caution though that other sub-solar metallicities have been reported
\citep{bailey_rv_2018}, and note that the cluster metallicity and
reddening are degenerate; if the reddening is lowered, the implied
metallicity will increase.  While a detailed redetermination of
these parameters from the Gaia photometry is beyond our scope, our
adopted values for both the reddening and metallicity are within the
range of previously reported values in the literature.

Overall, the data and models agree for masses above
$\approx$0.45\,$M_\odot$.  Below this mass, the data and models
diverge at colors redder than $\bpmrp\approx2.2$, in
the sense that the model isochrones have lower luminosities and bluer
colors than the observations.\footnote{The 100 Myr isochrone does
overlap well with the lower main sequence, but it fails to fit the
upper main sequence and red giants.}  The MIST isochrones showed a
comparable disagreement \citep{choi_mesa_2016}.  Proposed explanations
for the discrepancy between the models and observations have include
starspots and incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximum
sensitivity to age.

@@

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\subfloat{
% 			\includegraphics[width=0.49\textwidth]{f3a.pdf}
% 			\includegraphics[width=0.49\textwidth]{f3b.pdf}
% 		}
% 
%         \vspace{-0.5cm}
% 		\subfloat{
% 			\includegraphics[width=0.49\textwidth]{f3c.pdf}
% 			\includegraphics[width=0.49\textwidth]{f3d.pdf}
% 		}
% 	\end{center}
% 	\vspace{-0.7cm}
%   \caption{ {\bf TESS rotations periods versus color for \cn.} {\it
%   Top row}: The core and halo of \cn.  Sets $\mathcal{A}$ ({\it left})
%   and $\mathcal{B}$ ({\it right}) undergo successive stages of
%   cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
%   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
%   \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
%   Stars in the core and the halo of NGC\,2516 overlap with the
%   Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
%   members have longer rotation periods than Pleiades members.  {\it
%   Bottom row}: Binarity indicators in \cn.  Astrometric binaries
%   ($\mathrm{RUWE}>1.2$; 11\% of stars) appear in red; photometric
%   binaries ($>$0.3\,mag from an empirical isochrone; 20\% of stars)
%   are overplotted in orange.  Rotation periods for comparison field
%   stars are shown in Appendix~\ref{app:compstar}.
%   \label{fig:rot}
% 	}
% \end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3a.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for comparison field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3c.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for the rotators in
  Figure~\ref{fig:rot}.} Sets $\mathcal{A}$ ({\it top}) and
  $\mathcal{B}$ ({\it bottom}) undergo the same successive stages of
  cleaning.  Astrometric binaries ($\mathrm{RUWE}>1.2$; 11\% of stars)
  appear in red; photometric binaries ($>$0.3\,mag from an empirical
  isochrone; 20\% of stars) are overplotted in orange.  
  \label{fig:binarity}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
of $\approx$M2V.

@@

Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown on top and $\mathcal{B}$ on
the bottom.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.  The
``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrpo<1.2$.  The implications of this comparison for the age of
\cn\ are discussed in Section~\ref{disc:absage}.

Figure~\ref{fig:binarity} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr},
fixing the nodes by hand.  We then labeled any points brighter than
0.3\,mag above the cluster sequence as photometric binaries.  This
included 20\% of the overall cluster sample.  Many of the resulting
candidate binaries, though not all of them, appear below the slow
rotation sequence.  In set $\mathcal{B}$, over the color range where
the slow and fast sequence are present ($0.5<\bpmrpo<1.3$), there are
281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
281 stars into ``fast'' and ``slow'' sequences by eye,  the fraction
of stars showing signs of binarity in the slow and fast subsamples are
30\% (63/210) and 56\% (40/71), respectively.  We speculate on
possible explanations in Section~\ref{disc:rotn}.

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our
TESS pipeline succeeded in making light curves.  In other words, the
stars in the base sample are those for which we could have plausibly
detected a rotation period.  This color range is preferred because the
the slow sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.  With $\approx$3.5\,pc, the period detection rate of
61\% is relatively low, due to the combined effects of crowding and
saturation.

@@

    members have detected rotation periods consistent with a
